\textbf{ Mapping stellar surfaces\\ I: Degeneracies in the rotational light curve problem }

### 2021-01-27 paper1/ms.tex
more updates

\begin{abstract}
    %
    Thanks to missions like \emph{Kepler} and \emph{TESS}, we now
    have access to tens of thousands of high precision, fast
    cadence, and long baseline stellar photometric observations.
    %
    In principle, these light curves encode a vast amount of information about
    stellar variability and, in particular, about the distribution of
    starspots and other features on their surfaces.
    %
    Unfortunately, the problem of inferring stellar surface properties
    from a rotational light curve is famously ill-posed;
    any inference that relies exclusively on the light curve
    of a single star is \emph{necessarily} prior-dominated.
    %
    The goal of this paper is twofold:
    (1) to explore the various degeneracies affecting the stellar
    light curve ``inversion'' problem and their effect on
    what can and cannot be learned from a stellar surface
    given unresolved photometric measurements; and
    (2) to motivate ensemble analyses of the light curves of
    many stars at once as a powerful method of breaking
    these degeneracies.
    %
    %
    This is the first in a series of papers devoted to the development
    of novel algorithms and tools for the analysis of
    stellar light curves and spectral time series, with the
    explicit goal of enabling statistically robust inference
    about their surface properties.
\end{abstract}

@@

The goal of this paper is to explore the degeneracies at play in the
stellar surface mapping problem from a theoretical standpoint.
We will focus in particular on two sources of degeneracies: the
null space intrinsic to the mapping from a two-dimensional surface to
a one-dimensional light curve (\S\ref{sec:nullspace})
and the degeneracy due to the unknowability of the true baseline in
single-band photometry (\S\ref{sec:baseline}). Within each section we
will discuss ways to either break these degeneracies or explicitly marginalize
over degenerate quantities.
We will focus in particular on the power of ensemble analyses: the
joint analysis of many light curves of statistically ``similar''
stars. We will show that even though individual light curves are not very
constraining, light curves of many stars
observed at different inclinations can uniquely constrain certain
properties of the surfaces of those stars.
%
This idea was recently explored to some extent
in \citet{Morris2020}, who used ensemble analyses to derive constraints on
spot coverage areas as a function of stellar age. However, one of
the main conclusions of the present paper is that
quantities like \textbf{the
    total spot coverage, the typical spot contrast, and the total number of
    spots are not observables in single-band photometry.} Instead,
any constraints placed on these quantities, even in the context of
ensemble analyses, are driven by the choice of prior and other assumptions.

The present paper is also similar to \citet{Walkowicz2013} and
\citet{Basri2020}, who explored the information content of
stellar light curves from a large set of simulated spotted stellar
surfaces. While our paper is largely complementary to
that work, we instead approach the information content problem from
a purely theoretical---as opposed to empirical---point of view.
%
The present paper is the first in a series dedicated to the development of
techniques to perform robust inference about stellar surfaces from
unresolved photometric and spectroscopic measurements. The results
of this paper serve as the starting point for the development of
an interpretable Gaussian process for the ensemble analysis of
stellar light curves, which is the subject of
\citepalias[\citealt{PaperII}, hereafter][]{PaperII}.

@@

diagonal of $\mathbf{S}$ are the corresponding \emph{singular values}. If $\pmb{\mathcal{A}}$

@@

Note that because of the orthogonality of $\mathbf{V}$,
    \mathbf{V} \mathbf{V}^\top &= \mathbf{I}
    \mathbf{V}_\bullet \mathbf{V}_\bullet^\top  + \mathbf{V}_\circ \mathbf{V}_\circ^\top &= \mathbf{I}
    \mathbf{P} + \mathbf{N} &= \mathbf{I}
so it follows from Equations~(\ref{eq:yrow}) and (\ref{eq:ynull})
that $\mathbf{y} = \mathbf{y}_\bullet + \mathbf{y}_\circ$.

>>>>

\begin{abstract}
    %
    Thanks to missions like \emph{Kepler} and \emph{TESS}, we now
    have access to tens of thousands of high precision, fast
    cadence, and long baseline stellar photometric observations.
    %
    In principle, these light curves encode a vast amount of information about
    stellar variability and, in particular, about the distribution of
    starspots and other features on their surfaces.
    %
    Unfortunately, the problem of inferring stellar surface properties
    from a rotational light curve is famously ill-posed,
    as it often does not admit a unique solution.
    %
    Inference about the number, size, contrast, and location of spots
    can therefore depend very strongly on the assumptions of the model,
    the regularization scheme, or the prior.
    %
    The goal of this paper is twofold:
    (1) to explore the various degeneracies affecting the stellar
    light curve ``inversion'' problem and their effect on
    what can and cannot be learned from a stellar surface
    given unresolved photometric measurements; and
    (2) to motivate ensemble analyses of the light curves of
    many stars at once as a powerful data-driven alternative to common
    priors adopted in the literature.
    %
    We further derive novel results on the dependence of the null space
    on stellar inclination and limb darkening and
    show that single-band photometric measurements cannot uniquely
    constrain quantities like the spot contrast or the total
    spot coverage without the use of strong priors.
    %
    This is the first in a series of papers devoted to the development
    of novel algorithms and tools for the analysis of
    stellar light curves and spectral time series, with the
    explicit goal of enabling statistically robust inference
    about their surface properties.
\end{abstract}

@@

The goal of this paper is to explore the degeneracies at play in the
stellar surface mapping problem from a theoretical standpoint.
We will focus in particular on two sources of degeneracies: the
null space intrinsic to the mapping from a two-dimensional surface to
a one-dimensional light curve (\S\ref{sec:nullspace})
and the degeneracy due to the unknowability of the true baseline in
single-band photometry (\S\ref{sec:baseline}). Within each section we
will discuss ways to either break these degeneracies or marginalize
over unknowable quantities.
We will focus in particular on the power of ensemble analyses: the
joint analysis of many light curves of statistically ``similar''
stars. We will show that even though individual light curves are not very
constraining, light curves of many stars
observed at different inclinations can uniquely constrain certain
properties of the surfaces of those stars.
%
This idea was recently explored to some extent
in \citet{Morris2020}, who used ensemble analyses to derive constraints on
spot coverage areas as a function of stellar age. However, one of
the main conclusions of the present paper is that
quantities like \textbf{the
    total spot coverage, the typical spot contrast, and the total number of
    spots are not direct observables in single-band photometry.} Instead,
any constraints placed on these quantities, even in the context of
ensemble analyses, are usually driven by the choice of prior and other assumptions.

The present paper is also similar to \citet{Walkowicz2013} and
\citet{Basri2020}, who explored the information content of
stellar light curves from a large set of simulated spotted stellar
surfaces. While our paper is largely complementary to
that work, we instead approach the information content problem from
a theoretical---as opposed to empirical---point of view.
%
The present paper is the first in a series dedicated to the development of
techniques to perform robust inference about stellar surfaces from
unresolved photometric and spectroscopic measurements. The results
of this paper serve as the starting point for the development of
an interpretable Gaussian process for the ensemble analysis of
stellar light curves, which is the subject of
\citepalias[\citealt{PaperII}, hereafter][]{PaperII}.

@@

diagonal of $\mathbf{S}$ are the corresponding \emph{singular values}, arranged
in descending order. If $\pmb{\mathcal{A}}$

@@

Two things should be noted concerning this derivation. First,
because of the orthogonality of $\mathbf{V}$,
%
\begin{proof}{test_PN}
    \mathbf{V} \mathbf{V}^\top &= \mathbf{I}
    \nonumber                                                 \\
    \mathbf{V}_\bullet \mathbf{V}_\bullet^\top  + \mathbf{V}_\circ \mathbf{V}_\circ^\top &= \mathbf{I}
    \nonumber \\
    \mathbf{P} + \mathbf{N} &= \mathbf{I}
    \quad,
\end{proof}
%
so it follows from Equations~(\ref{eq:yrow}) and (\ref{eq:ynull})
that $\mathbf{y} = \mathbf{y}_\bullet + \mathbf{y}_\circ$.

Second, the singular value decomposition is not always unique. The
matrix of singular values $\mathbf{S}$ is unique provided we arrange
them in decreasing order, but if there are degenerate or zero-valued
singular values (as is the case here), the matrices $\mathbf{U}$ and
$\mathbf{V}^\top$ (and thus $\mathbf{V}^\top_\bullet$ and $\mathbf{V}^\top_\circ$)
are only well-defined up to arbitrary unitary
transformations. However, the quantities
$\mathbf{V}_\bullet\mathbf{V}^\top_\bullet$
and
$\mathbf{V}_\circ\mathbf{V}^\top_\circ$
(which define the operators $\mathbf{P}$ and $\mathbf{N}$ above) \emph{are}
unique, so the decomposition into null space and preimage is always well-defined.
